---
title: "Data from radiation and weather in Viikki field"
subtitle: "2017-2024"
author: "Pedro J. Aphalo"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
# library(tint)
# invalidate cache when the package version changes
knitr::opts_chunk$set(tidy = FALSE) # , cache.extra = packageVersion('tint')
options(htmltools.dir.version = FALSE)
```

When new data is available, first run 'read-data.Rmd', possibly after clearing the 'knitr' cache.

```{r}
library(tidyverse)
library(ggridges)
library(lubridate)

options(warnPartialMatchAttr = FALSE,
        warnPartialMatchDollar = FALSE,
        warnPartialMatchArgs = FALSE)
```


```{r}
rm(list = ls(pattern = "*"))
load("data-rda/hour_2015_latest.tb.rda")
range(hour_2015_latest.tb$time)
```

## Important note

ggplot stores the whole data frame in the ggplot object. Here we have 50 variables and many thousands of observations. I have changed the code to first select the columns to be used before building the plot. This reduces the size of plots from 72 MB to 2-3 MB and also makes plotting a lot faster.

## Density plots by month (ridge plots)

## PAR photon irradiance

One hour averages for sun above the horizon of measurements with LI-190 quantum sensor every 5 seconds. During winter-time some snow could have accumulated on the sensor.


```{r}
hour_2015_latest.tb %>%
  filter(PAR_umol > 0.01) %>%
  select(month_name, PAR_umol, calendar_year) %>%
  ggplot(aes(x = PAR_umol, y = month_name)) +
  geom_density_ridges(na.rm = TRUE, scale = 2.5, fill = "darkgreen",
                      alpha = 0.33, bandwidth = 0.05) +
  scale_x_log10(breaks = c(0.01, 0.03, 0.1, 0.3, 1, 3, 10, 30, 100, 300, 1000, 3000), 
                minor_breaks = NULL, 
                labels = scales::format_format(digits = 3, scientific = FALSE)) +
  scale_y_discrete(drop=FALSE) +
#  coord_cartesian(xlim = c(1, 3000)) +
  labs(x = expression(PAR~~photon~~irradiance~~(mu*mol~m^{-2}~s^{-1})), y = "Month of the year") -> par_month_rigdes.fig
par_month_rigdes.fig
```

```{r}
par_month_rigdes.fig +
  facet_wrap(~calendar_year) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))

```


```{r}
pdf("figures-pdf/par-dens-month-year-fig.pdf", width = 10, height = 6)
print(par_month_rigdes.fig)
dev.off()
```

## Fraction of PAR photon irradiance that is diffuse

One minute averages for sun above the horizon of measurements with LI-190 quantum sensor every 5 seconds. During winter-time some snow could have accumulated on the sensor.

```{r}
# hour_2015_2022.tb %>%
#   filter(PAR_umol > 5) %>%
#   select(month_name, PAR_diff_fr) %>%
#   ggplot(aes(x = PAR_diff_fr, y = month_name)) +
#   geom_vline(xintercept = c(0.1, 1), linetype = "dashed", colour = "red") +
#   geom_density_ridges(na.rm = TRUE, scale = 4.5, fill = "grey50", alpha = 0.33, bandwidth = 0.01) +
#   scale_y_discrete(drop=FALSE) +
#   scale_x_continuous(breaks = c(0, 0.1, 0.25, 0.5, 0.75, 1), limits = c(0, 1.3)) +
#   labs(x = "Diffuse fraction of total PAR", y = "Month of the year") -> diff_month_rigdes.fig
# diff_month_rigdes.fig
```

```{r}
# pdf("figures-pdf/diff-par-dens-month-year-fig.pdf", width = 10, height = 6)
# print(diff_month_rigdes.fig)
# dev.off()
```

```{r}
hour_2015_latest.tb %>%
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) %>%
  select(month_name, air_DP) %>%
  ggplot(aes(x = air_DP, y = month_name)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "red") +
  geom_density_ridges(na.rm = TRUE, scale = 3.5, fill = "grey50", alpha = 0.33) +
  scale_y_discrete(drop=FALSE) +
#  scale_x_continuous(breaks = c(0, 0.1, 0.25, 0.5, 0.75, 1)) +
  labs(x = "Dew point (C)", y = "Month of the year") -> dew_point_month_rigdes.fig
dew_point_month_rigdes.fig
```

```{r}
hour_2015_latest.tb %>%
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) %>%
  select(month_name, air_temp_C) %>%
  ggplot(aes(x = air_temp_C, y = month_name)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "red") +
  geom_density_ridges(na.rm = TRUE, scale = 3.5, fill = "grey50", alpha = 0.33) +
  scale_y_discrete(drop=FALSE) +
#  scale_x_continuous(breaks = c(0, 0.1, 0.25, 0.5, 0.75, 1)) +
  labs(x = "Air temperature (C)", y = "Month of the year") -> air_temp_month_rigdes.fig
air_temp_month_rigdes.fig
```

```{r}
hour_2015_latest.tb %>%
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) %>%
  select(month_name, air_RH) %>%
  ggplot(aes(x = air_RH, y = month_name)) +
  geom_vline(xintercept = c(0, 100), linetype = "dashed", colour = "red") +
  geom_density_ridges(na.rm = TRUE, scale = 3.5, fill = "grey50", alpha = 0.33) +
  scale_y_discrete(drop=FALSE) +
#  scale_x_continuous(breaks = c(0, 0.1, 0.25, 0.5, 0.75, 1)) +
  labs(x = "Air relative humidity (%)", y = "Month of the year") -> air_temp_month_rigdes.fig
air_temp_month_rigdes.fig
```

```{r}
hour_2015_latest.tb %>%
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) %>%
  select(month_name, wind_speed) %>%
  ggplot(aes(x = wind_speed, y = month_name)) +
  geom_density_ridges(na.rm = TRUE, scale = 1.5, fill = "grey50", alpha = 0.33) +
  scale_y_discrete(drop=FALSE) +
#  scale_x_continuous(breaks = c(0, 0.1, 0.25, 0.5, 0.75, 1)) +
  labs(x = "Wind speed (m s-1)", y = "Month of the year") -> wind_speed_month_rigdes.fig
wind_speed_month_rigdes.fig
```

```{r}
hour_2015_latest.tb %>%
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) %>%
  select(month_name, air_pressure) %>%
  ggplot(aes(x = air_pressure, y = month_name)) +
  geom_density_ridges(na.rm = TRUE, scale = 1.5, fill = "grey50", alpha = 0.33) +
  scale_y_discrete(drop=FALSE) +
#  scale_x_continuous(breaks = c(0, 0.1, 0.25, 0.5, 0.75, 1)) +
  labs(x = "Air Pressure (Pa)", y = "Month of the year") -> 
air_pressure_month_rigdes.fig
air_pressure_month_rigdes.fig
```

```{r}
load("data-rda/hour_soil_2020_latest.tb.rda")
names(hour_soil_2020_latest.tb)
range(hour_soil_2020_latest.tb$time)
```

```{r}
hour_soil_2020_latest.tb %>%
  group_by(time) %>%
  transmute(time,
            PAR_umol,
#            PAR_diff_fr,
            global_watt,
            wind_speed,
            wind_direction,
            air_temp_C,
            air_RH,
            air_DP,
            air_pressure,
            day_of_year,
            month_of_year,
            month_name,
            calendar_year,
            solar_time,
#            was_sunny,
            T_5cm = median(T_5cm_1, T_5cm_2, T_5cm_3, na.rm = TRUE),
            T_10cm = median(T_10cm_1, T_10cm_2, T_10cm_3, na.rm = TRUE),
            T_20cm = median(T_20cm_1, T_20cm_2, T_20cm_3, na.rm = TRUE),
            T_30cm = median(T_30cm_1, T_30cm_2, T_30cm_3, na.rm = TRUE),
            T_40cm = median(T_40cm_1, T_40cm_2, T_40cm_3, na.rm = TRUE),
            T_50cm = median(T_50cm_1, T_50cm_2, T_50cm_3, na.rm = TRUE),
            VWC_5cm = median(VWC_5cm_1, VWC_5cm_2, VWC_5cm_3, na.rm = TRUE),
            VWC_10cm = median(VWC_10cm_1, VWC_10cm_2, VWC_10cm_3, na.rm = TRUE),
            VWC_20cm = median(VWC_20cm_1, VWC_20cm_2, VWC_20cm_3, na.rm = TRUE),
            VWC_30cm = median(VWC_30cm_1, VWC_30cm_2, VWC_30cm_3, na.rm = TRUE),
            VWC_40cm = median(VWC_40cm_1, VWC_40cm_2, VWC_40cm_3, na.rm = TRUE),
            VWC_50cm = median(VWC_50cm_1, VWC_50cm_2, VWC_50cm_3, na.rm = TRUE),
            Ka_5cm = median(Ka_5cm_1, Ka_5cm_2, Ka_5cm_3, na.rm = TRUE),
            Ka_10cm = median(Ka_10cm_1, Ka_10cm_2, Ka_10cm_3, na.rm = TRUE),
            Ka_20cm = median(Ka_20cm_1, Ka_20cm_2, Ka_20cm_3, na.rm = TRUE),
            Ka_30cm = median(Ka_30cm_1, Ka_30cm_2, Ka_30cm_3, na.rm = TRUE),
            Ka_40cm = median(Ka_40cm_1, Ka_40cm_2, Ka_40cm_3, na.rm = TRUE),
            Ka_50cm = median(Ka_50cm_1, Ka_50cm_2, Ka_50cm_3, na.rm = TRUE),
            BulkEC_5cm = median(BulkEC_5cm_1, BulkEC_5cm_2, BulkEC_5cm_3, na.rm = TRUE),
            BulkEC_10cm = median(BulkEC_10cm_1, BulkEC_10cm_2, BulkEC_10cm_3, na.rm = TRUE),
            BulkEC_20cm = median(BulkEC_20cm_1, BulkEC_20cm_2, BulkEC_20cm_3, na.rm = TRUE),
            BulkEC_30cm = median(BulkEC_30cm_1, BulkEC_30cm_2, BulkEC_30cm_3, na.rm = TRUE),
            BulkEC_40cm = median(BulkEC_40cm_1, BulkEC_40cm_2, BulkEC_40cm_3, na.rm = TRUE),
            BulkEC_50cm = median(BulkEC_50cm_1, BulkEC_50cm_2, BulkEC_50cm_3, na.rm = TRUE),
            VWC_20cm_x = median(VWC_1, VWC_2, VWC_3, na.rm = TRUE),
            T_20cm_x = median(T_1, T_2, T_3, na.rm = TRUE)
  ) -> hour_soil.tb
names(hour_soil.tb)
```

```{r}
hour_soil.tb %>%
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) %>%
  select(time, T_5cm, T_10cm, T_20cm, T_30cm, T_40cm, T_50cm) %>%
  ggplot() +
  geom_line(aes(y = T_5cm, x = time), color = "red") +
  geom_line(aes(y = T_10cm, x = time)) +
  geom_line(aes(y = T_20cm, x = time), color = "darkgreen") +
  geom_line(aes(y = T_30cm, x = time), color = "darkblue") +
  geom_line(aes(y = T_50cm, x = time), color = "darkred") +
#  scale_x_continuous(breaks = c(0, 0.1, 0.25, 0.5, 0.75, 1)) +
  labs(y = "Soil temperature at 5 to 50 cm depth (C)", x = "Time") -> 
soil_t_50cm.fig
soil_t_50cm.fig
```

```{r}
hour_soil.tb %>%
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) %>%
  select(time, VWC_5cm, VWC_10cm, VWC_20cm, VWC_30cm, VWC_50cm) %>%
  ggplot() +
  geom_line(aes(y = VWC_5cm, x = time), color = "red") +
  geom_line(aes(y = VWC_10cm, x = time)) +
  geom_line(aes(y = VWC_20cm, x = time), color = "darkgreen") +
  geom_line(aes(y = VWC_30cm, x = time), color = "darkblue") +
  geom_line(aes(y = VWC_50cm, x = time), color = "darkred") +
#  scale_x_continuous(breaks = c(0, 0.1, 0.25, 0.5, 0.75, 1)) +
  labs(y = "Soil water v/v ()", x = "Time") +
  expand_limits(y = 0) -> 
soil_t_50cm.fig
soil_t_50cm.fig
```
```{r}
hour_soil.tb %>%
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) %>%
  select(time, BulkEC_10cm, BulkEC_20cm, BulkEC_30cm, BulkEC_50cm) %>%
  ggplot() +
  geom_line(aes(y = BulkEC_10cm, x = time)) +
  geom_line(aes(y = BulkEC_20cm, x = time), color = "darkgreen") +
  geom_line(aes(y = BulkEC_30cm, x = time), color = "darkblue") +
  geom_line(aes(y = BulkEC_50cm, x = time), color = "darkred") +
#  scale_x_continuous(breaks = c(0, 0.1, 0.25, 0.5, 0.75, 1)) +
  labs(y = "Bulk EC (dS/m)", x = "Time") ->
soil_t_50cm.fig
soil_t_50cm.fig
```

```{r}
hour_soil.tb %>%
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) %>%
  select(time, Ka_10cm, Ka_20cm, Ka_30cm, Ka_50cm) %>%
  ggplot() +
#  geom_line(aes(y = Ka_10cm, x = time)) +
  geom_line(aes(y = Ka_20cm, x = time), color = "darkgreen") +
  geom_line(aes(y = Ka_30cm, x = time), color = "darkblue") +
  geom_line(aes(y = Ka_50cm, x = time), color = "darkred") +
#  scale_x_continuous(breaks = c(0, 0.1, 0.25, 0.5, 0.75, 1)) +
  labs(y = "Relative Dielectric Permittivity (1..80)", x = "Time") +
  expand_limits(y = 0) -> 
soil_t_50cm.fig
soil_t_50cm.fig
```
```{r}
hour_soil.tb %>%
  select(time, T_20cm_x) %>%
  ggplot(aes(time, T_20cm_x)) +
  geom_line() +
  labs(y = "Soil temperature at 20 cm (C)", x = "Time") -> 
soil_t_20cm.fig
soil_t_20cm.fig
```

```{r}
hour_soil.tb %>%
  select(time, VWC_20cm_x) %>%
  ggplot(aes(time, VWC_20cm_x)) +
  geom_line() +
  labs(y = "Soil water at 20 cm v/v ()", x = "Time") +
  expand_limits(y = 0) -> 
soil_wvc_20cm.fig
soil_wvc_20cm.fig
```
