---
title: "Data from radiation and weather in Viikki field"
subtitle: "One year of data 2017-2023"
author: "Pedro J. Aphalo"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(tint)
# invalidate cache when the package version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tint'))
options(htmltools.dir.version = FALSE)
```

When new data is available, first run 'read-data.Rmd', possibly after clearing the 'knitr' cache.

```{r}
library(tidyverse)
library(ggridges)
library(lubridate)

options(warnPartialMatchAttr = FALSE,
        warnPartialMatchDollar = FALSE,
        warnPartialMatchArgs = FALSE)
```


```{r}
rm(list = ls(pattern = "*"))
load("data-rda/minute_2015_2023.tb.rda")
range(minute_2015_2023.tb$time)
```

Radiation data are not usable for the period November to March as snow intermitently accumulates on the sensors input optics. 
```{r}
good_irrad <- function(x) {x[x$month_of_year %in% 4:10 & !is.na(x$was_sunny), ]}
minute_all.tb <- minute_2015_2023.tb
minute_rad.tb<- good_irrad(minute_all.tb)
```

## Important note

ggplot stores the whole data frame in the ggplot object. Here we have 50 variables and many thousands of observations. I have changed the code to first select the columns to be used before building the plot. This reduces the size of plots from 72 MB to 2-3 MB and also makes plotting a lot faster.

## Plots by sun position

```{r}
minute_rad.tb %>%
  select(sun_elevation, red_far_red) %>%
  ggplot(aes(sun_elevation, red_far_red)) +
  geom_smooth(na.rm = TRUE) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(title = paste("R:FR ratio in Viikki,", as.Date(min(minute_rad.tb$time)), "to", as.Date(max(minute_rad.tb$time))),
       x = "Solar elevation angle (degrees)",
       y = "R:FR, red to far-red photon ratio") +
  annotate(geom = "text", x = 42, y = 1.07, hjust = 0,
           label = paste("n =", nrow(minute_rad.tb))) +
  theme_bw() -> rfr_sun_elev.fig
rfr_sun_elev.fig
```

```{r}
pdf("figures-pdf/r-fr-sun-elev-year-fig.pdf", width = 8, height = 6)
print(rfr_sun_elev.fig)
dev.off()
```

```{r}
minute_rad.tb %>%
  select(sun_azimuth, red_far_red) %>%
  ggplot(aes(sun_azimuth, red_far_red)) +
  geom_smooth(na.rm = TRUE) +
  scale_x_continuous(breaks = c(90, 180, 270)) +
  #  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(title = paste("R:FR ratio in Viikki,", as.Date(min(minute_rad.tb$time)), "to", as.Date(max(minute_rad.tb$time))),
       x = "Solar azimuth angle (degrees)",
       y = "R:FR, red to far-red photon ratio") +
  annotate(geom = "text", x = 180, y = 1.1,
           label = paste("n =", nrow(subset(minute_2015_2023.tb, !is.na(was_sunny))))) +
  theme_bw() -> rfr_sun_azimuth.fig
rfr_sun_azimuth.fig
```

```{r}
pdf("figures-pdf/r-fr-sun-azimuth-year-fig.pdf", width = 8, height = 6)
print(rfr_sun_azimuth.fig)
dev.off()
```

```{r}
minute_rad.tb %>%
  select(sun_elevation, PAR_umol) %>%
  ggplot(aes(sun_elevation, PAR_umol)) +
  geom_density2d() +
  geom_smooth(na.rm = TRUE) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(title = paste("R:FR ratio in Viikki,", as.Date(min(minute_rad.tb$time)), "to", as.Date(max(minute_rad.tb$time))),
       x = "Solar elevation angle (degrees)",
       y = "PAR") +
  annotate(geom = "text", x = 42, y = 1.07, hjust = 0,
           label = paste("n =", nrow(minute_rad.tb))) +
  expand_limits(y = -150) +
  theme_bw() -> par_sun_elev.fig
par_sun_elev.fig
```

## Density plots by month (ridge plots)

### R:FR

One minute averages for sun above the horizon of measurements with Skye sensor every 5 seconds. During winter-time some snow could have accumulated on the sensor.

```{r}
minute_rad.tb %>%
  filter(sun_elevation > 0) %>%
  select(month_name, red_far_red, calendar_year) %>%
  ggplot(aes(x = red_far_red, y = month_name)) +
  geom_vline(xintercept = 1.16, linetype = "dashed", colour = "red") +
  # geom_density_ridges(na.rm = TRUE, fill = "darkred", alpha = 0.3,
  #                     quantile_lines = TRUE, quantiles = 2) +
  geom_density_ridges(na.rm = TRUE, fill = "darkred", alpha = 0.3) +
  scale_y_discrete(drop=FALSE) +
  xlim(c(0.5, 1.7)) +
  labs(x = "R:FR photon ratio", y = "Month of the year") -> rfr_month_rigdes.fig
rfr_month_rigdes.fig
```

```{r}
rfr_month_rigdes.fig +
  facet_wrap(~calendar_year)

```

```{r}
pdf("figures-pdf/r-fr-dens-month-year-fig.pdf", width = 10, height = 6)
print(rfr_month_rigdes.fig)
dev.off()
```

One minute averages including values for sun below the horizon

```{r, eval=FALSE}
minute_2015_2023.tb %>%
  filter(sun_elevation > -10) %>%  # > -10 does not work
  select(month_name, red_far_red, calendar_year) %>%
  na.omit() %>%
  ggplot(aes(x = red_far_red, y = month_name)) +
  geom_vline(xintercept = 1.16, linetype = "dashed", colour = "red") +
  # geom_density_ridges(na.rm = TRUE, fill = "darkred", alpha = 0.3,
  #                     quantile_lines = TRUE, quantiles = 2) +
  geom_density_ridges(na.rm = TRUE, fill = "darkred", alpha = 0.3) +
  scale_y_discrete(drop=FALSE) +
  xlim(c(0.5, 1.7)) +
  labs(x = "R:FR photon ratio", y = "Month of the year") -> rfr_month_rigdes_all.fig
rfr_month_rigdes_all.fig
```

```{r, eval=FALSE}
rfr_month_rigdes_all.fig +
  facet_wrap(~calendar_year)
```
## PAR photon irradiance

One minute averages for sun above the horizon of measurements with LI-190 quantum sensor every 5 seconds. During winter-time some snow could have accumulated on the sensor.


```{r}
minute_2015_2023.tb %>%
  filter(sun_elevation > 0) %>%
  select(month_name, PAR_umol, calendar_year) %>%
  ggplot(aes(x = PAR_umol, y = month_name)) +
  geom_vline(xintercept = 2000, linetype = "dotted") +
  geom_density_ridges(na.rm = TRUE, scale = 2.5, fill = "darkgreen", alpha = 0.33) +
  scale_x_log10(breaks = c(0.03, 0.1, 0.3, 1, 3, 10, 30, 100, 300, 1000, 3000), 
                minor_breaks = NULL, 
                labels = scales::format_format(digits = 3, scientific = FALSE)) +
  scale_y_discrete(drop=FALSE) +
  coord_cartesian(xlim = c(0.03, 3000)) +
  labs(x = expression(PAR~~photon~~irradiance~~(mu*mol~m^{-2}~s^{-1})), y = "Month of the year") -> par_month_rigdes.fig
par_month_rigdes.fig
```

```{r}
par_month_rigdes.fig +
  facet_wrap(~calendar_year) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))

```


```{r}
pdf("figures-pdf/par-dens-month-year-fig.pdf", width = 10, height = 6)
print(par_month_rigdes.fig)
dev.off()
```

## Fraction of PAR photon irradiance that is diffuse

One minute averages for sun above the horizon of measurements with LI-190 quantum sensor every 5 seconds. During winter-time some snow could have accumulated on the sensor.

```{r}
  minute_rad.tb %>%
  filter(sun_elevation > 10 & PAR_umol > 5) %>%
  select(month_name, PAR_diff_fr) %>%
  ggplot(aes(x = PAR_diff_fr, y = month_name)) +
  geom_vline(xintercept = c(0.1, 1), linetype = "dashed", colour = "red") +
  geom_density_ridges(na.rm = TRUE, scale = 4.5, fill = "grey50", alpha = 0.33) +
  scale_y_discrete(drop=FALSE) +
  scale_x_continuous(breaks = c(0, 0.1, 0.25, 0.5, 0.75, 1)) +
  labs(x = "Diffuse fraction of total PAR", y = "Month of the year") -> diff_month_rigdes.fig
diff_month_rigdes.fig
```

```{r}
pdf("figures-pdf/diff-par-dens-month-year-fig.pdf", width = 10, height = 6)
print(diff_month_rigdes.fig)
dev.off()
```

```{r}
minute_2015_2023.tb %>%
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) %>%
  select(month_name, air_DP) %>%
  ggplot(aes(x = air_DP, y = month_name)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "red") +
  geom_density_ridges(na.rm = TRUE, scale = 3.5, fill = "grey50", alpha = 0.33, 
                      bandwidth = .9) +
  scale_y_discrete(drop=FALSE) +
#  scale_x_continuous(breaks = c(0, 0.1, 0.25, 0.5, 0.75, 1)) +
  labs(x = "Dew point (C)", y = "Month of the year") -> dew_point_month_rigdes.fig
dew_point_month_rigdes.fig
```

```{r}
minute_2015_2023.tb %>%
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) %>%
  select(month_name, air_temp_C) %>%
  ggplot(aes(x = air_temp_C, y = month_name)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "red") +
  geom_density_ridges(na.rm = TRUE, scale = 3.5, fill = "grey50", alpha = 0.33,
                      bandwidth = 0.9) +
  scale_y_discrete(drop=FALSE) +
#  scale_x_continuous(breaks = c(0, 0.1, 0.25, 0.5, 0.75, 1)) +
  labs(x = "Air temperature (C)", y = "Month of the year") -> air_temp_month_rigdes.fig
air_temp_month_rigdes.fig
```

```{r}
minute_2015_2023.tb %>%
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) %>%
  select(month_name, air_RH) %>%
  ggplot(aes(x = air_RH, y = month_name)) +
  geom_vline(xintercept = c(0, 100), linetype = "dashed", colour = "red") +
  geom_density_ridges(na.rm = TRUE, scale = 3.5, fill = "grey50", alpha = 0.33) +
  scale_y_discrete(drop=FALSE) +
#  scale_x_continuous(breaks = c(0, 0.1, 0.25, 0.5, 0.75, 1)) +
  labs(x = "Air relative humidity (%)", y = "Month of the year") -> air_temp_month_rigdes.fig
air_temp_month_rigdes.fig
```

```{r}
minute_2015_2023.tb %>%
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) %>%
  select(month_name, wind_speed) %>%
  ggplot(aes(x = wind_speed, y = month_name)) +
  geom_density_ridges(na.rm = TRUE, scale = 1.5, fill = "grey50", alpha = 0.33) +
  scale_y_discrete(drop=FALSE) +
#  scale_x_continuous(breaks = c(0, 0.1, 0.25, 0.5, 0.75, 1)) +
  labs(x = "Wind speed (m s-1)", y = "Month of the year") -> wind_speed_month_rigdes.fig
wind_speed_month_rigdes.fig
```

```{r}
minute_2015_2023.tb %>%
#  filter(sun_elevation > 0 & PAR_Den_Avg > 5) %>%
  select(month_name, air_pressure) %>%
  ggplot(aes(x = air_pressure, y = month_name)) +
  geom_density_ridges(na.rm = TRUE, scale = 1.5, fill = "grey50", alpha = 0.33) +
  scale_y_discrete(drop=FALSE) +
#  scale_x_continuous(breaks = c(0, 0.1, 0.25, 0.5, 0.75, 1)) +
  labs(x = "Air Pressure (Pa)", y = "Month of the year") -> 
  air_pressure_month_rigdes.fig

air_pressure_month_rigdes.fig
```
