---
title: "PAR diffuse fraction under clear sky"
subtitle: "Data simulated with the TUV model"
author: "Pedro J. Aphalo"
date: "2024-08-27"
output: html_document
---

```{r setup, include=FALSE}
rm(list = ls(pattern = "*"))
knitr::opts_chunk$set(tidy = FALSE, cache = FALSE)
```

When new data is available, first run 'read-data.Rmd', possibly after clearing the 'knitr' cache.

```{r}
library(dplyr)
library(ggplot2)
library(lubridate)
library(photobiology)
library(photobiologyInOut)
library(photobiologyWavebands)

theme_set(theme_bw())

options(warnPartialMatchAttr = FALSE,
        warnPartialMatchDollar = FALSE,
        warnPartialMatchArgs = FALSE)
```

## Diffuse UV-B, UV-A2, UV-A1, Blue, Green, Red and Far-red fraction from TUV

Read Quick TUV output files for different solar zenith angles and split the
components of total irradiance into separate spectra. We select files with
`low` in their names, with low levels of aerosols that seem match better the
data as can be seen in the next section. One can change `"low"` by `"4str"`
for the default aerosols from Quick TUV.

```{r, message = FALSE, warning = FALSE}

files <- list.files(pattern = "tuv", full.names = TRUE)

SZA.mspct <- source_mspct()

for (f in files) {
  name <- gsub("tuv-default-|-4str.htm|-low.htm", "", basename(f))
  print(name)
  SZA.mspct[[name]] <- read_qtuv_txt(f, ozone.du = 300)
}

SZA_tot.mspct <- source_mspct()
SZA_diff.mspct <- source_mspct()
SZA_dir.mspct <- source_mspct()
for (n in names(SZA.mspct)) {
  SZA_tot.mspct[[n]] <-
    SZA.mspct[[n]][ , c("w.length", "s.e.irrad")]
  SZA_diff.mspct[[n]] <-
    with(SZA.mspct[[n]], source_spct(w.length = w.length, s.e.irrad = s.e.irrad.diff.down))
  SZA_dir.mspct[[n]] <-
    with(SZA.mspct[[n]], source_spct(w.length = w.length, s.e.irrad = s.e.irrad.dir))
}
```

Compute irradiances and the diffuse fraction for different SZAs.

```{r}
wavebands <- c(list(PAR = PAR()), Plant_bands("sensory"))
wbands_tot.tb <- 
  q_irrad(SZA_tot.mspct, wavebands, scale.factor = 1e6)
wbands_dir.tb <- 
  q_irrad(SZA_dir.mspct, wavebands, scale.factor = 1e6)
wbands_diff.tb <- 
  q_irrad(SZA_diff.mspct, wavebands, scale.factor = 1e6)

wbands.tb <-
  data.frame(
    SZA = as.numeric(gsub("tuv-|\\.htm", "", wbands_tot.tb$spct.idx)),
    SEA = 90 - as.numeric(gsub("tuv-|\\.htm", "", wbands_tot.tb$spct.idx))
  )

for (col in setdiff(colnames(wbands_tot.tb), "spct.idx")) {
  wbands.tb[[paste(col, "tot", sep = ".")]] <- wbands_tot.tb[[col]]
  wbands.tb[[paste(col, "dir", sep = ".")]] <- wbands_dir.tb[[col]]
  wbands.tb[[paste(col, "diff", sep = ".")]] <- wbands_diff.tb[[col]]
  wbands.tb[[paste(col, "diff_fr", sep = ".")]] <- 
    wbands_diff.tb[[col]] / wbands_tot.tb[[col]]
}

colnames(wbands.tb) <- 
  gsub("Q_]|Q_|\\.ISO|\\.CIE|\\.Sellaro|\\.Smith20", "", colnames(wbands.tb))

colnames(wbands.tb)
```

Plot to check that data look good. We use solar elevation angle for the plot.

```{r}
ggplot(wbands.tb, aes(SEA, PAR.diff_fr)) +
  geom_line() +
  expand_limits(y = 0)
```

```{r}
# We fit a splines obtaining functions that can be used to obtain by interpolation
# estimates of the diffuse fraction for any solar elevation covered by the
# simulations.

spl_funs.ls <- list()
diff_fr_cols <- grep("diff[_]fr", colnames(wbands.tb), value = TRUE)

for (col in diff_fr_cols) {
  spl_funs.ls[[col]] <- splinefun(wbands.tb[["SEA"]], wbands.tb[[col]])
}
```

```{r}
ggplot(wbands.tb, aes(SEA, PAR.diff_fr)) +
  stat_function(fun = spl_funs.ls[["PAR.diff_fr"]], xlim = c(-10, 90), colour = "black") +
  geom_point(na.rm = TRUE, colour =  "orange") +
  scale_x_continuous(name = "Solar elevation angle (degrees)",
                     breaks = c(-10, 0, 15, 30, 45, 60, 75, 90)) +
  labs(y = expression("Diffuse fraction, "*Q[s] / Q[t])) +
  expand_limits(y = 0)
```

```{r}
saveRDS(spl_funs.ls, file = "diff-fr-spl-funs.RDS")
```

